\vspace{24pt} Comparison of Different Data Layouts for Columnar In-Memory Databases on the Basis of an Example Application

### 2018-08-30 paper/paper.tex
kv format

\usepackage{hyperref}
\newcommand{\quotes}[1]{``{#1}''}

@@

This approach avoids using timestamps, which is a major advantage in term of memory consumption. Also querying a time series is supposed to be faster compared to the frame format. However, the different positions within the frames have to be calculated, which can be a tedious thing. Due to the use of frames the data is normalized by time on the one hand. On the other hand, there is a loss of information. Furthermore, there could be the possibility of missing points. In this case, Wang et al.~\cite{wang} calculate those points via SED distance. We refrain from that. The overall frame format table with a framegroup size of ten minutes and a frame size of 15 seconds consumes circa 441 MB of main memory, when stored in column layout.

@@

A third approach we evaluated is the key-value format.
\subsection{Optimizations}

this layout is splitting the sample sample points into frame groups of a fixed length with a fixed frequency of data points. We decided to use a frame size of 15 seconds, since this is the median frequency the data was recorded (see \ref{sec:ds}). Furthermore, we use a frame group size of ten minutes.

@@

\section{Conclusion}
% \bibliographystyle{ACM-Reference-Format}
\bibliographystyle{plain}
\bibliography{references}

>>>>

\usepackage{hyperref}
\usepackage{listings}

\newcommand{\quotes}[1]{``{#1}''}

\usepackage{xcolor}

\colorlet{punct}{red!60!black}
\definecolor{background}{HTML}{EEEEEE}
\definecolor{delim}{RGB}{20,105,176}
\colorlet{numb}{magenta!60!black}

% There is no lst-language highlighting json code nicely. Hence, we define it here.
\lstdefinelanguage{json}{
    basicstyle=\normalfont\ttfamily,
    numbers=left,
    numberstyle=\scriptsize,
    stepnumber=1,
    numbersep=8pt,
    showstringspaces=false,
    breaklines=true,
    frame=lines,
    backgroundcolor=\color{background},
    literate=
     *{0}{{{\color{numb}0}}}{1}
      {1}{{{\color{numb}1}}}{1}
      {2}{{{\color{numb}2}}}{1}
      {3}{{{\color{numb}3}}}{1}
      {4}{{{\color{numb}4}}}{1}
      {5}{{{\color{numb}5}}}{1}
      {6}{{{\color{numb}6}}}{1}
      {7}{{{\color{numb}7}}}{1}
      {8}{{{\color{numb}8}}}{1}
      {9}{{{\color{numb}9}}}{1}
      {:}{{{\color{punct}{:}}}}{1}
      {,}{{{\color{punct}{,}}}}{1}
      {\{}{{{\color{delim}{\{}}}}{1}
      {\}}{{{\color{delim}{\}}}}}{1}
      {[}{{{\color{delim}{[}}}}{1}
      {]}{{{\color{delim}{]}}}}{1},
}

@@

This approach avoids using timestamps, which is a major advantage in term of memory consumption. Also querying a time series is supposed to be faster compared to the frame format. However, the different positions within the frames have to be calculated, which can be a tedious thing. Due to the use of frames the data is normalized by time on the one hand. On the other hand, there is a loss of information. Furthermore, there could be the possibility of missing points. In this case, Wang et al.~\cite{wang} calculate those points via SED distance. We refrain from that. The overall frame format table with a framegroup size of ten minutes and a frame size of 15 seconds consumes circa 441 MB of main memory, when stored in column layout. In total, there are more than 2 million rows.

@@

A third storage layout we evaluated is the key-value format. Therefore, we decided to encode a whole trajectory as an array of arrays containing longitude, latitude and the corresponding timestamp. Since an array cannot be stored in the SAP Hana database, we take its string representation and filed it in a column of data type \textit{NCLOB}. An example can be found in code listing \ref{lst:json}. We do not normalize by time. Hence, there is no loss in terms of GPS data. Besides this information, there are the trajectory identifier, start and end time and the mininum bounding rectangle present in the table. In total, there are 10,594 (equivalent to the amount of trajectories after data cleansing) consuming about 650 MB of main memory in column store.

\lstset{language=json}
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}[frame=single, firstnumber=1, caption=Key-Value format example entry,label=lst:json]
"[
    [
        "0001-01-01 00:00:03",
        22.512432,
        113.913933
    ],
    [
        "0001-01-01 00:00:58",
        22.5163,
        113.917068
    ],
    ...
]"
\end{lstlisting}
\end{minipage}

\subsection{Optimizations}

this layout is splitting the sample sample points into frame groups of a fixed length with a fixed frequency of data points. We decided to use a frame size of 15 seconds, since this is the median frequency the data was recorded (see section \ref{sec:ds}). Furthermore, we use a frame group size of ten minutes.

@@

\section{Conclusion}

% \bibliographystyle{ACM-Reference-Format}
\bibliographystyle{plain}
\bibliography{references}
